---
title: "Shandong Data Analysis - OB & GYN only"
author: "Yao He"
date: "09/17/2021"
output:
  pdf_document: default
---

# SET UP

```{r, include = FALSE}
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(warning=FALSE)
options(scipen=999)

```

```{r,include=FALSE,message=FALSE,warning=FALSE}
# Load packages
library(knitr)
library(plyr)
library(readr)
library(dplyr) # mutate(), rename()
library(ggplot2)
library(lattice)
library(nlme)
library(lme4)
library(lspline)
library(stargazer)
library(UWbe536)
library(glarma)

#library(kableExtra)
#library(geepack)
#library(miceadds)
#library(data.table)
#library(doBy)
#library(tidyr)
library(sandwich)
library(MASS)
library(lmtest)
library(cowplot)
library(tseries) #adf.test Dickey-Fuller test; p<0.05, reject null that time series is not stationary

setwd("~/Documents/R/Research/Shandong/Data/")

```

```{r goldentheme,include=F}
# Create the golden plot template
goldentheme <- theme(
  panel.background = element_rect(fill = "white"),
  aspect.ratio = ((1 + sqrt(5))/2)^(-1),
  axis.ticks.length = unit(0.5, "char"),
  axis.line.x.top = element_line(size = 0.2),
  axis.line.x.bottom = element_line(size = 0.2),
  axis.ticks.x = element_line(size = 0.2),
  axis.text.x = element_text(color = "black", size = 12),
  axis.title.x = element_text(size = 12,
                              margin = margin(t = 7.5, r = 0, b = 0, l = 0)),
  axis.ticks.y = element_blank(),
  axis.text.y = element_text(color = "black", size = 12,
                             margin = margin(t = 0, r = -4, b = 0, l = 0)),
  axis.title.y = element_text(size = 12,
                              margin = margin(t = 0, r = 7.5, b = 0, l = 0)),
  legend.key = element_rect(fill = NA, color = NA),
  panel.grid.major.x = element_blank(),
  panel.grid.major.y = element_line(color = "gray45", size = 0.2),
  strip.background = element_blank(),
  strip.text.x = element_text(size=12),
  strip.text.y = element_blank(), 
  strip.placement = "outside", 
  panel.spacing.x = unit(1.25, "lines"),
  panel.spacing.y = unit(1, "lines") 
)

class(goldentheme)

goldentheme_facet <- theme(
  panel.background = element_rect(fill = "white"),
  axis.ticks.length = unit(0.5, "char"),
  axis.line.x.top = element_line(size = 0.2),
  axis.line.x.bottom = element_line(size = 0.2),
  axis.ticks.x = element_line(size = 0.2),
  axis.text.x = element_text(color = "black", size = 12),
  axis.title.x = element_text(size = 12,
                              margin = margin(t = 7.5, r = 0, b = 0, l = 0)),
  axis.ticks.y = element_blank(),
  axis.text.y = element_text(color = "black", size = 12,
                             margin = margin(t = 0, r = -4, b = 0, l = 0)),
  axis.title.y = element_text(size = 12,
                              margin = margin(t = 0, r = 7.5, b = 0, l = 0)),
  legend.key = element_rect(fill = NA, color = NA),
  panel.grid.major.x = element_blank(),
  panel.grid.major.y = element_line(color = "gray45", size = 0.2),
  strip.background = element_blank(),
  strip.text.x = element_text(size=12),
  strip.text.y = element_blank(), 
  strip.placement = "outside", 
  panel.spacing.x = unit(1.25, "lines"),
  panel.spacing.y = unit(1, "lines") 
)

class(goldentheme_facet)

```

```{r}
# create a function to replace any space in x axis labels with a new line
addline_format <- function(x,...){
    gsub('\\s','\n',x)
}

```


# OBSTETRICS

## Outpatient

### Data & Descriptive

```{r}
# read in data
datOS <- read_csv("Data/monthly_outpatient_surgery_inpatient.csv")

# subset for outpatient OBSTETRICS
datObOUT <- subset(datOS, grepl("产", datOS$keshi) & datOS$division=="outpatient")

# combine counts if a monthly count was divided among more than 1 subdivisions of the obstetrics department
datObOUT1 <- datObOUT %>% group_by(year, month) %>% summarise_at(vars(count), sum)

# create Time var: first ever month in dataset = 0
datObOUT1$time <- (datObOUT1$year-2015)*12+(datObOUT1$month-1)

# create Time1 var: 0 centered at Jan 2020 (1 time point before the pandemic actually happened in Feb 2020)
datObOUT1$time1 <- datObOUT1$time - 60

# create COVID1 var to indicate just Feb 2020 to make the drop in that month pop.
datObOUT1$COVID1 <- with(datObOUT1, ifelse(year == 2020 & month == 2, 1, 0))

# create COVID2 var to indicate the time points starting from Mar 2020. 
datObOUT1$COVID2 <- with(datObOUT1, ifelse(time1 >= 2, 1, 0))

# create Post var; 1 should start at April 2020
datObOUT1$post <- with(datObOUT1, ifelse(time1 >= 3, (year-2020)*12+(month-3), 0))

```

```{r}
# initial chart of # cases by month; just raw data (2015-2020)
# ggplot(data = datObOUT1, aes(x = time, y = count)) +
#       geom_point() +
#       geom_line() +
#       ggtitle("Obstetric outpatient visits (2015-2021 May)") +
#       labs(x = "Time", y = "Number of obstetric outpatient visits") +
#       scale_x_continuous(breaks=c(0,12,24,36,48,60,72,76),
#                        labels=addline_format(c("Jan 2015","Jan 2016","Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","May 2021"))) +
#       geom_vline(xintercept = 61, linetype = "dotted") +
#     annotate(geom="text", x=68, y=25000, label="COVID-19",size=5) +
#           expand_limits(y = 0) +
#       goldentheme +
#   theme(axis.text.x = element_text(size = 8.5)) -> pp.ObOUT.15
# pp.ObOUT.15

```

```{r}
# initial chart of # cases by month; just raw data (2017-2020)
ggplot(data = datObOUT1, aes(x = time, y = count)) +
      geom_point() +
      geom_line() +
      ggtitle("Obstetric outpatient visits (Jan 2017-May 2021)") +
      labs(x = "Time", y = "Number of obstetric outpatient visits") +
      scale_x_continuous(limits=c(24,76), breaks=c(24,36,48,60,72,76),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","May 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=66.5, y=25000, label="COVID-19",size=5) +
          expand_limits(y = 0) +
      goldentheme +
  theme(axis.text.x = element_text(size = 9.5)) -> pp.ObOUT.17
pp.ObOUT.17

```

### Analysis

$$ ln(E(y_t)) = \alpha_0 + \sum_{m=2}^{12} \beta_m Month + \beta_1 Time_t + \beta_2 COVID1 + \beta_3 COVID2_t + \beta_4 Post_t $$

```{r}
# adjust for monthly seasonality, January as the control，no offset for newborn population becuase 2021 data are not accurate
mod.ObOUT <- glm.nb(count ~ time1 + COVID1 + COVID2 + post 
                  #  + offset(log(newborn))
                  #  +feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec
                  + as.factor(month), 
                 data = subset(datObOUT1, year >= 2017))

#summary(mod.ObOUT)
lincom(mod.ObOUT)

# Extract the fitted values from this fitted model
datObOUT1$fit <- NA
datObOUT1$fit[datObOUT1$year>=2017] <- fitted(mod.ObOUT)

# Use AR(1)
coef.ObOUT <- coeftest(mod.ObOUT, vcov=NeweyWest(mod.ObOUT,lag=1,prewhite = F))

# put coef, CI, and p-value for time1, COVID, post together
m.ObOUT <- matrix(, nrow = 4, ncol = 4)
m.ObOUT[,1]<-exp(coef.ObOUT[c(2:5),1])
m.ObOUT[,2]<-exp(coef.ObOUT[c(2:5),1]-1.96*coef.ObOUT[c(2:5),2])
m.ObOUT[,3]<-exp(coef.ObOUT[c(2:5),1]+1.96*coef.ObOUT[c(2:5),2])
m.ObOUT[,4]<-coef.ObOUT[c(2:5),4]
rownames(m.ObOUT) <- c("time1", "COVID1","COVID2", "post")
colnames(m.ObOUT) <- c("exp(beta)", "95%low", "95%up", "p-value")
round(m.ObOUT, 5)

```

Compared to January 2020, the number of obstetric outpatient visits in February 2020, when the pandemic and lockdown first started, was `r round((1-m.ObOUT[2,1])*100,1)`% lower (95% CI `r round(m.ObOUT[2,2],3)`, `r m.ObOUT(m.OBm.4[2,3],3)`). The decrease was statistically significant (p < 0.0001). 

```{r}
# Counterfactual or expected Y values based on the original model
X0.ObOUT <- X.ObOUT <- model.matrix(mod.ObOUT)
X0.ObOUT[,3:5] <- 0
datObOUT1$exp[datObOUT1$year>=2017] <- as.vector(exp(X0.ObOUT%*%coef.ObOUT)) # save expected Y values to the monthly dataset

```

```{r}
####### 10,000 simulations
Sigma.ObOUT <- NeweyWest(mod.ObOUT,lag=1,prewhite = F)
set.seed(12)
beta.ObOUT <- mvrnorm(n = 10000, coef.ObOUT[,1], Sigma.ObOUT)

D.ObOUT <- Y_sum.ObOUT <- Y0_sum.ObOUT <- D_P.ObOUT <- RR.ObOUT <- NA

for (i in 1:10000){
  Y_sum.ObOUT[i] <- sum(as.vector(exp(X.ObOUT%*%beta.ObOUT[i,]))[38:53])
  Y0_sum.ObOUT[i] <- sum(as.vector(exp(X0.ObOUT%*%beta.ObOUT[i,]))[38:53])
  D.ObOUT[i] <- Y0_sum.ObOUT[i] - Y_sum.ObOUT[i]
  D_P.ObOUT[i] <- (Y0_sum.ObOUT[i] - Y_sum.ObOUT[i]) / Y0_sum.ObOUT[i]
  RR.ObOUT[i] <- Y_sum.ObOUT[i] / Y0_sum.ObOUT[i]
}

# Fitted post
mean(Y_sum.ObOUT)
quantile(Y_sum.ObOUT, c(0.025,0.975)) #95% CI

# Counterfactual post
mean(Y0_sum.ObOUT)
quantile(Y0_sum.ObOUT, c(0.025,0.975)) 

# !!! Lost visits (observed - counterfactual/forecasted)
mean(D.ObOUT) 
quantile(D.ObOUT, c(0.025,0.975))
p.D.ObOUT <- length(D.ObOUT[D.ObOUT < 0]) / length(D.ObOUT) # p-value; if p=0, write it as p<0.0001
p.D.ObOUT

mean(D_P.ObOUT)  ## percentage of decrease
quantile(D_P.ObOUT, c(0.025,0.975))
p.D_P.ObOUT <- length(D_P.ObOUT[D_P.ObOUT < 0]) / length(D_P.ObOUT) # p-value
p.D_P.ObOUT

mean(RR.ObOUT)  ## RR
quantile(RR.ObOUT, c(0.025,0.975))

```

```{r}
### Chart for the model analysis
# CI around the line, just for the fitted line.
#agg_all$fit <- predict(m0,type="response")
datObOUT1$fit_up[datObOUT1$year>=2017] <- exp(predict(mod.ObOUT) + 1.96*predict(mod.ObOUT, se.fit=T)$se.fit)
datObOUT1$fit_low[datObOUT1$year>=2017] <- exp(predict(mod.ObOUT) - 1.96*predict(mod.ObOUT,se.fit=T)$se.fit)

# chart of observed versus expected
ggplot(data = datObOUT1, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp, colour = "Expected")) +
        geom_line(aes(y = fit, colour = "Fitted")) + 
      geom_point(aes(y = count, colour = "Fitted")) +
      ggtitle("Obstetric outpatient visits: Observed versus expected (Jan 2017 - May 2021)") +
      labs(x = "Time", y = "Obstetric outpatient visits", color = "") +
      scale_x_continuous(limits=c(24,82), breaks=c(24,36,48,60,72,82),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Nov 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=67, y=11000, label="COVID-19",size=5) +
          expand_limits(y = c(8000:26000)) +
      goldentheme +
  theme(axis.text.x = element_text(size = 9.5)) -> pp.ObOUT.ove
pp.ObOUT.ove

```


## Surgery

### Data & Descriptive

```{r}
# subset for surgery OBSTETRICS
datObSUR <- subset(datOS, grepl("产", datOS$keshi) & datOS$division=="surgery")

# combine counts if a monthly count was divided among more than 1 subdivisions of the obstetrics department
datObSUR1 <- datObSUR %>% group_by(year, month) %>% summarise_at(vars(count), sum)

# create Time var: first ever month in dataset = 0
datObSUR1$time <- (datObSUR1$year-2015)*12+(datObSUR1$month-1)

# create Time1 var: 0 centered at Jan 2020 (1 time point before the pandemic actually happened in Feb 2020)
datObSUR1$time1 <- datObSUR1$time - 60

# create COVID1 var to indicate just Feb 2020 to make the drop in that month pop.
datObSUR1$COVID1 <- with(datObSUR1, ifelse(year == 2020 & month == 2, 1, 0))

# create COVID2 var to indicate the time points starting from Mar 2020. 
datObSUR1$COVID2 <- with(datObSUR1, ifelse(time1 >= 2, 1, 0))

# create Post var; 1 should start at April 2020
datObSUR1$post <- with(datObSUR1, ifelse(time1 >= 3, (year-2020)*12+(month-3), 0))

```

```{r}
# initial chart of # cases by month; just raw data (2015-2020)
# ggplot(data = datObSUR1, aes(x = time, y = count)) +
#       geom_point() +
#       geom_line() +
#       ggtitle("Obstetric surgeries (2015-2021 May)") +
#       labs(x = "Time", y = "Number of obstetric surgeries") +
#       scale_x_continuous(breaks=c(0,12,24,36,48,60,72,76),
#                        labels=addline_format(c("Jan 2015","Jan 2016","Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","May 2021"))) +
#       geom_vline(xintercept = 61, linetype = "dotted") +
#     annotate(geom="text", x=68, y=1200, label="COVID-19",size=5) +
#           expand_limits(y = 0) +
#       goldentheme +
#   theme(axis.text.x = element_text(size = 8.5)) -> pp.ObSUR.15
# pp.ObSUR.15

```

```{r}
# initial chart of # cases by month; just raw data (2017-2020)
ggplot(data = datObSUR1, aes(x = time, y = count)) +
      geom_point() +
      geom_line() +
      ggtitle("Obstetric surgeries (Jan 2017-May 2021)") +
      labs(x = "Time", y = "Number of obstetric surgeries") +
      scale_x_continuous(limits=c(24,76), breaks=c(24,36,48,60,72,76),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","May 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=66.5, y=1200, label="COVID-19",size=5) +
          expand_limits(y = 0) +
      goldentheme +
  theme(axis.text.x = element_text(size = 9.5)) -> pp.ObSUR.17
pp.ObSUR.17

```

### Analysis

$$ ln(E(y_t)) = \alpha_0 + \sum_{m=2}^{12} \beta_m Month + \beta_1 Time_t + \beta_2 COVID1 + \beta_3 COVID2_t + \beta_4 Post_t $$

```{r}
# adjust for monthly seasonality, January as the control，no offset for newborn population becuase 2021 data are not accurate
mod.ObSUR <- glm.nb(count ~ time1 + COVID1 + COVID2 + post 
                  #  + offset(log(newborn))
                  #  +feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec
                  + as.factor(month), 
                 data = subset(datObSUR1, year >= 2017))

#summary(mod.ObOUT)
lincom(mod.ObSUR)

# Extract the fitted values from this fitted model
datObSUR1$fit <- NA
datObSUR1$fit[datObSUR1$year>=2017] <- fitted(mod.ObSUR)

# Use AR(1)
coef.ObSUR <- coeftest(mod.ObSUR, vcov=NeweyWest(mod.ObSUR,lag=1,prewhite = F))

# put coef, CI, and p-value for time1, COVID, post together
m.ObSUR <- matrix(, nrow = 4, ncol = 4)
m.ObSUR[,1]<-exp(coef.ObSUR[c(2:5),1])
m.ObSUR[,2]<-exp(coef.ObSUR[c(2:5),1]-1.96*coef.ObSUR[c(2:5),2])
m.ObSUR[,3]<-exp(coef.ObSUR[c(2:5),1]+1.96*coef.ObSUR[c(2:5),2])
m.ObSUR[,4]<-coef.ObSUR[c(2:5),4]
rownames(m.ObSUR) <- c("time1", "COVID1","COVID2", "post")
colnames(m.ObSUR) <- c("exp(beta)", "95%low", "95%up", "p-value")
round(m.ObSUR, 5)

```

Compared to January 2020, the number of obstetric surgeries in February 2020, when the pandemic and lockdown first started, was `r round((1-m.ObOUT[2,1])*100,1)`% lower (95% CI `r round(m.ObOUT[2,2],3)`, `r m.ObOUT(m.OBm.4[2,3],3)`). The decrease was statistically significant (p < 0.0001). 

```{r}
# Counterfactual or expected Y values based on the original model
X0.ObSUR <- X.ObSUR <- model.matrix(mod.ObSUR)
X0.ObSUR[,3:5] <- 0
datObSUR1$exp[datObSUR1$year>=2017] <- as.vector(exp(X0.ObSUR%*%coef.ObSUR)) # save expected Y values to the monthly dataset

```

```{r}
####### 10,000 simulations
Sigma.ObSUR <- NeweyWest(mod.ObSUR,lag=1,prewhite=F)
set.seed(12)
beta.ObSUR <- mvrnorm(n = 10000, coef.ObSUR[,1], Sigma.ObSUR)

D.ObSUR <- Y_sum.ObSUR <- Y0_sum.ObSUR <- D_P.ObSUR <- RR.ObSUR <- NA

for (i in 1:10000){
  Y_sum.ObSUR[i] <- sum(as.vector(exp(X.ObSUR%*%beta.ObSUR[i,]))[38:53])
  Y0_sum.ObSUR[i] <- sum(as.vector(exp(X0.ObSUR%*%beta.ObSUR[i,]))[38:53])
  D.ObSUR[i] <- Y0_sum.ObSUR[i] - Y_sum.ObSUR[i]
  D_P.ObSUR[i] <- (Y0_sum.ObSUR[i] - Y_sum.ObSUR[i]) / Y0_sum.ObSUR[i]
  RR.ObSUR[i] <- Y_sum.ObSUR[i] / Y0_sum.ObSUR[i]
}

# Fitted post
mean(Y_sum.ObSUR)
quantile(Y_sum.ObSUR, c(0.025,0.975)) #95% CI

# Counterfactual post
mean(Y0_sum.ObSUR)
quantile(Y0_sum.ObSUR, c(0.025,0.975)) 

# !!! Lost visits (observed - counterfactual/forecasted)
mean(D.ObSUR) 
quantile(D.ObSUR, c(0.025,0.975))
p.D.ObSUR <- length(D.ObSUR[D.ObSUR < 0]) / length(D.ObSUR) # p-value; if p=0, write it as p<0.0001
p.D.ObSUR

mean(D_P.ObSUR)  ## percentage of decrease
quantile(D_P.ObSUR, c(0.025,0.975))
p.D_P.ObSUR <- length(D_P.ObSUR[D_P.ObSUR < 0]) / length(D_P.ObSUR) # p-value
p.D_P.ObSUR

mean(RR.ObSUR)  ## RR
quantile(RR.ObSUR, c(0.025,0.975))

```

```{r}
### Chart for the model analysis
# CI around the line, just for the fitted line.
#agg_all$fit <- predict(m0,type="response")
datObSUR1$fit_up[datObSUR1$year>=2017] <- exp(predict(mod.ObSUR) + 1.96*predict(mod.ObSUR, se.fit=T)$se.fit)
datObSUR1$fit_low[datObSUR1$year>=2017] <- exp(predict(mod.ObSUR) - 1.96*predict(mod.ObSUR,se.fit=T)$se.fit)

# chart of observed versus expected
ggplot(data = datObSUR1, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp, colour = "Expected")) +
        geom_line(aes(y = fit, colour = "Fitted")) + 
      geom_point(aes(y = count, colour = "Fitted")) +
      ggtitle("Obstetric surgeries: Observed versus expected (Jan 2017 - May 2021)") +
      labs(x = "Time", y = "Obstetric surgeries", color = "") +
      scale_x_continuous(limits=c(24,82), breaks=c(24,36,48,60,72,82),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Nov 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=67, y=1050, label="COVID-19",size=5) +
          expand_limits(y = c(500:1200)) +
      goldentheme +
  theme(axis.text.x = element_text(size = 9.5)) -> pp.ObSUR.ove
pp.ObSUR.ove

```

## Inpatient

### Data & Descriptive

```{r, echo=T}
# read in data
datOB <- read_csv("Data/allob.csv")

# create Time var: first ever month in dataset = 0
datOB$time <- (datOB$year-2015)*12+(datOB$month-1)

# create Time1 var: 0 centered at Jan 2020 (1 time point before the pandemic actually happened in Feb 2020)
datOB$time1 <- datOB$time - 60

# create COVID1 var to indicate just Feb 2020 to make the drop in that month pop.
datOB$COVID1 <- with(datOB, ifelse(year == 2020 & month == 2, 1, 0))

# create COVID2 var to indicate the time points starting from Mar 2020. 
datOB$COVID2 <- with(datOB, ifelse(time1 >= 2, 1, 0))

# create Post var; 1 should start at April 2020
datOB$post <- with(datOB, ifelse(time1 >= 3, (year-2020)*12+(month-3), 0))

# change wrong age data
datOB$age <- with(datOB, ifelse(age==1000, 0, age))

# exclude newborns (age==0) from the data
datOBp <- subset(datOB, age>0)

```

```{r}
# Summarize monthly counts
datOBp %>% 
  group_by(time, year, month, time1, COVID1, COVID2, post) %>% 
  count(time) -> datObIN

# june and july 2021 are very low because of opening new sub-hospital
# impute with average of may and aug 2021
datObIN$n <- with(datObIN, ifelse(year==2021 & month==6 | year==2021 & month==7,
                                  round((datObIN$n[datObIN$year==2021 & datObIN$month==5] +
                                         datObIN$n[datObIN$year==2021 & datObIN$month==8]) / 2, 0), 
                                  n))

```


```{r}
# initial chart of # cases by month; just raw data (2017-2021)
ggplot(data = datObIN, aes(x = time, y = n)) +
      geom_point() +
      geom_line() +
      ggtitle("Obstetric inpatient admissions (Jan 2017-May 2021)") +
      labs(x = "Time", y = "Number of obstetric admissions") +
      scale_x_continuous(limits=c(24,76), breaks=c(24,36,48,60,72,76),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","May 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=66.5, y=2200, label="COVID-19",size=5) +
          expand_limits(y = 0) +
      goldentheme +
  theme(axis.text.x = element_text(size = 9.5)) -> pp.ObIN.17
pp.ObIN.17

```

### Analysis

```{r}
# create yearly sine cosine trend values - Ended up not being used
#datOBm$sinyr <- with(datOBm, sin((2*pi*time1)/12))
#datOBm$cosyr <- with(datOBm, cos((2*pi*time1)/12))

# create month indicator for adjusting for each month - ENDED UP NOT BEING USED
#datOBm$feb <- with(datOBm, ifelse(month==2,1,0))
#datOBm$mar <- with(datOBm, ifelse(month==3,1,0))
# datOBm$apr <- with(datOBm, ifelse(month==4,1,0))
# datOBm$may <- with(datOBm, ifelse(month==5,1,0))
# datOBm$jun <- with(datOBm, ifelse(month==6,1,0))
# datOBm$jul <- with(datOBm, ifelse(month==7,1,0))
# datOBm$aug <- with(datOBm, ifelse(month==8,1,0))
# datOBm$sep <- with(datOBm, ifelse(month==9,1,0))
# datOBm$oct <- with(datOBm, ifelse(month==10,1,0))
# datOBm$nov <- with(datOBm, ifelse(month==11,1,0))
# datOBm$dec <- with(datOBm, ifelse(month==12,1,0))

```


$$ ln(E(y_t)) = \alpha_0 + \sum_{m=2}^{12} \beta_m Month + \beta_1 Time_t + \beta_2 COVID1 + \beta_3 COVID2_t + \beta_4 Post_t $$

```{r}
# adjust for monthly seasonality, January as the control，no offset for newborn population becuase 2021 data are not accurate
mod.ObIN <- glm.nb(n ~ time1 + COVID1 + COVID2 + post 
                  #  + offset(log(newborn))
                  #  +feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec
                  + as.factor(month), 
                 data = subset(datObIN, year >= 2017))

#summary(mod.OBm)
lincom(mod.ObIN)

# Extract the fitted values from this fitted model
datObIN$fit <- NA
datObIN$fit[datObIN$year>=2017] <- fitted(mod.ObIN)

# Use AR(1)
coef.ObIN <- coeftest(mod.ObIN, vcov=NeweyWest(mod.ObIN,lag=1,prewhite = F))

# put coef, CI, and p-value for time1, COVID, post together
m.ObIN <- matrix(, nrow = 4, ncol = 4)
m.ObIN[,1]<-exp(coef.ObIN[c(2:5),1])
m.ObIN[,2]<-exp(coef.ObIN[c(2:5),1]-1.96*coef.ObIN[c(2:5),2])
m.ObIN[,3]<-exp(coef.ObIN[c(2:5),1]+1.96*coef.ObIN[c(2:5),2])
m.ObIN[,4]<-coef.ObIN[c(2:5),4]
rownames(m.ObIN) <- c("time1", "COVID1","COVID2", "post")
colnames(m.ObIN) <- c("exp(beta)", "95%low", "95%up", "p-value")
round(m.ObIN, 5)

```

Compared to January 2020, the number of obstetric inpatient visits in February 2020, when the pandemic and lockdown first started, was `r round((1-m.ObIN[2,1])*100,1)`% lower (95% CI `r round(m.ObIN[2,2],3)`, `r round(m.ObIN[2,3],3)`). The decrease was statistically significant (p < 0.0001). 

After the pandemic started, the number of obstetric inpatient visits started to recover, but the monthly increase was only `r round((m.ObIN[1,1]-m.ObIN[3,1])*100,1)`% (95% CI `r round(m.ObIN[1,3]-m.ObIN[3,3],3)`, `r round(m.ObIN[1,2]-m.ObIN[3,2],3)`). This monthly increase was not statistically significant (p = 0.157).


```{r}
# Counterfactual or expected Y values based on the original model
X0.ObIN <- X.ObIN <- model.matrix(mod.ObIN)
X0.ObIN[,3:5] <- 0
datObIN$exp[datObIN$year>=2017] <- as.vector(exp(X0.ObIN%*%coef.ObIN)) # save expected Y values to the monthly dataset

```

```{r}
####### 10,000 simulations
Sigma.ObIN <- NeweyWest(mod.ObIN,lag=1,prewhite=F)
set.seed(12)
beta.ObIN <- mvrnorm(n = 10000, coef.ObIN[,1], Sigma.ObIN)

D.ObIN <- Y_sum.ObIN <- Y0_sum.ObIN <- D_P.ObIN <- RR.ObIN <- NA

for (i in 1:10000){
  Y_sum.ObIN[i] <- sum(as.vector(exp(X.ObIN%*%beta.ObIN[i,]))[38:53])
  Y0_sum.ObIN[i] <- sum(as.vector(exp(X0.ObIN%*%beta.ObIN[i,]))[38:53])
  D.ObIN[i] <- Y0_sum.ObIN[i] - Y_sum.ObIN[i]
  D_P.ObIN[i] <- (Y0_sum.ObIN[i] - Y_sum.ObIN[i]) / Y0_sum.ObIN[i]
  RR.ObIN[i] <- Y_sum.ObIN[i] / Y0_sum.ObIN[i]
}

# Fitted post
mean(Y_sum.ObIN)
quantile(Y_sum.ObIN, c(0.025,0.975)) #95% CI

# Counterfactual post
mean(Y0_sum.ObIN)
quantile(Y0_sum.ObIN, c(0.025,0.975)) 

# !!! Lost visits (observed - counterfactual/forecasted)
mean(D.ObIN) 
quantile(D.ObIN, c(0.025,0.975))
p.D.ObIN <- length(D.ObIN[D.ObIN < 0]) / length(D.ObIN) # p-value; if p=0, write it as p<0.0001
p.D.ObIN

mean(D_P.ObIN)  ## percentage of decrease
quantile(D_P.ObIN, c(0.025,0.975))
p.D_P.ObIN <- length(D_P.ObIN[D_P.ObIN < 0]) / length(D_P.ObIN) # p-value
p.D_P.ObIN

mean(RR.ObIN)  ## RR
quantile(RR.ObIN, c(0.025,0.975))

```

```{r}
### Chart for the model analysis
# CI around the line, just for the fitted line.
#agg_all$fit <- predict(m0,type="response")
datObIN$fit_up[datObIN$year>=2017] <- exp(predict(mod.ObIN) + 1.96*predict(mod.ObIN, se.fit=T)$se.fit)
datObIN$fit_low[datObIN$year>=2017] <- exp(predict(mod.ObIN) - 1.96*predict(mod.ObIN, se.fit=T)$se.fit)

# chart of observed versus expected
ggplot(data = datObIN, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp, colour = "Expected")) +
        geom_line(aes(y = fit, colour = "Fitted")) + 
      geom_point(aes(y = n, colour = "Fitted")) +
      ggtitle("Obstetric inpatient admissions: Observed vs. expected (Jan 2017 - Dec 2021)") +
      labs(x = "Time", y = "Obstetric admissions", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Dec 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=67, y=1900, label="COVID-19",size=5) +
          expand_limits(y = c(1000:2000)) +
      goldentheme +
  theme(axis.text.x = element_text(size = 9.5)) -> pp.ObIN.ove
pp.ObIN.ove


```

```{r}
#really lag 1?

library(tseries)

## outpatient
adf.test(datObOUT1$count) # larger than 0.05, dp not reject the null that the time series is not stationary. Since it is not stationary, it means we might not need to use autoregressive models.

# What's the lag?
acf(datObOUT1$count) #sort of decaying slowly, suitable for AR
pacf(datObOUT1$count) #seems lag = 2 is the last significant spike

## surgery
adf.test(datObSUR1$count) # larger than 0.05, dp not reject the null that the time series is not stationary. Since it is not stationary, it means we might not need to use autoregressive models.

# What's the lag?
acf(datObSUR1$count) #sort of decaying slowly, suitable for AR
pacf(datObSUR1$count) #seems lag = 2 is the last significant spike

## inpatient data
adf.test(datObIN$n) # larger than 0.05, dp not reject the null that the time series is not stationary. Since it is not stationary, it means we might not need to use autoregressive models.

# What's the lag?
acf(datObIN$n) #sort of decaying slowly, suitable for AR
pacf(datObIN$n) #seems lag = 1 is the last significant spike

# so lag really is 1

```

### Plot all 3 charts together

```{r}
# outpatient
ggplot(data = datObOUT1, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp), color = "#0072B2", linetype = "dashed") +
        geom_line(aes(y = fit), color = "#0072B2",) + 
      geom_point(aes(y = count), color = "#0072B2", size = 1) +
      ggtitle("A. Outpatient visits") +
      labs(x = "", y = "", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                         labels=addline_format(c("","","","","",""))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
          expand_limits(y = c(8000:26000)) +
       scale_y_continuous(trans="log10") +
      goldentheme_facet +
  theme(axis.title.y=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none",
        plot.title = element_text(face = "bold")) -> pp.ObOUT.ove1
pp.ObOUT.ove1

# surgeries
ggplot(data = datObSUR1, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp), color = "#000000", linetype = "dashed") +
        geom_line(aes(y = fit), color = "#000000") + 
      geom_point(aes(y = count), color = "#000000", size = 1) +
      ggtitle("B. Surgeries") +
      labs(x = "", y = "", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                         labels=addline_format(c("","","","","",""))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
          expand_limits(y = c(500:1200)) +
         scale_y_continuous(trans="log10") +
      goldentheme_facet +
  theme(axis.title.y=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none",
        plot.title = element_text(face = "bold")) -> pp.ObSUR.ove1
pp.ObSUR.ove1

#inpatient
ggplot(data = datObIN, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp), color = "#D55E00", linetype = "dashed") +
        geom_line(aes(y = fit), color = "#D55E00") + 
      geom_point(aes(y = n), color = "#D55E00", size = 1) +
      ggtitle("C. Inpatient admissions") +
      labs(x = "Time", y = "", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Dec 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=67.5, y=1900, label="COVID-19",size=4) +
          expand_limits(y = c(1000:2000)) +
         scale_y_continuous(trans="log10") +
      goldentheme_facet +
  theme(axis.text.x = element_text(size = 10),
        axis.title.y=element_blank(),
        legend.position = "none",
        plot.title = element_text(face = "bold")) -> pp.ObIN.ove1
pp.ObIN.ove1

```

```{r}
#dev.new(width=5, height=10, noRStudioGD = TRUE)

threez <- plot_grid(pp.ObOUT.ove1, pp.ObSUR.ove1, pp.ObIN.ove1, nrow = 3,
          rel_heights = c(1, 1, 1.15))

ggsave("~/Documents/RESEARCH/Shandong/Manuscripts/Figures/Ob3in1.pdf", width = 5, height = 8)

threez

```

```{r}
# all 3 lines in one chart
ggplot() +
        geom_vline(xintercept = 61, linetype = "dotted") +
  geom_ribbon(data = datObOUT1, aes(x = time, ymin = fit_low, ymax = fit_up), fill = "#0072B2", alpha = 0.15) +
  geom_ribbon(data = datObSUR1, aes(x = time, ymin = fit_low, ymax = fit_up), fill = "#000000", alpha = 0.15) +
  geom_ribbon(data = datObIN, aes(x = time, ymin = fit_low, ymax = fit_up), fill = "#D55E00", alpha = 0.15) +
      geom_line(data = datObOUT1, aes(x = time, y = exp), color = "#0072B2", linetype = "dashed") +
      geom_line(data = datObSUR1, aes(x = time, y = exp), color = "#000000", linetype = "dashed") +
      geom_line(data = datObIN, aes(x = time, y = exp), color = "#D55E00", linetype = "dashed") +
        geom_line(data = datObOUT1, aes(x = time, y = fit), color = "#0072B2",) + 
        geom_line(data = datObSUR1, aes(x = time, y = fit), color = "#000000") + 
        geom_line(data = datObIN, aes(x = time, y = fit), color = "#D55E00") + 
      geom_point(data = datObOUT1, aes(x = time, y = count), color = "#0072B2", size = 1) +
      geom_point(data = datObSUR1, aes(x = time, y = count), color = "#000000", size = 1) +
      geom_point(data = datObIN, aes(x = time, y = n), color = "#D55E00", size = 1) +
      ggtitle("") +
      labs(x = "Time", y = "", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Dec 2021"))) +
      annotate(geom="text", x=67, y=3500, label="COVID-19",size=5) +
          annotate(geom="text", x=29, y=15000, label="Outpatient visits",size=4, color = "#0072B2") +
        annotate(geom="text", x=30.5, y=2400, label="Inpatient admissions",size=4, color = "#D55E00") +
          annotate(geom="text", x=27, y=800, label="Surgeries",size=4, color = "#000000") +
              expand_limits(y = c(100:30000)) +
       scale_y_continuous(trans="log10") +
      goldentheme +
  theme(axis.title.y=element_blank(),
        legend.position = "none",
        plot.title = element_blank()) -> pp.Ob.all.ove

ggsave("~/Documents/RESEARCH/Shandong/Manuscripts/Figures/Ob3in1_v2.pdf", width = 8, height = 5)

pp.Ob.all.ove

```


***

# GYNECOLOGY

## Outpatient

### Data & Descriptive

```{r}
# subset for outpatient GYNO
datGynOUT <- subset(datOS, grepl("妇", datOS$keshi) & datOS$division=="outpatient")

# combine counts if a monthly count was divided among more than 1 subdivisions of the obstetrics department
datGynOUT1 <- datGynOUT %>% group_by(year, month) %>% summarise_at(vars(count), sum)

# create Time var: first ever month in dataset = 0
datGynOUT1$time <- (datGynOUT1$year-2015)*12+(datGynOUT1$month-1)

# create Time1 var: 0 centered at Jan 2020 (1 time point before the pandemic actually happened in Feb 2020)
datGynOUT1$time1 <- datGynOUT1$time - 60

# create COVID1 var to indicate just Feb 2020 to make the drop in that month pop.
datGynOUT1$COVID1 <- with(datGynOUT1, ifelse(year == 2020 & month == 2, 1, 0))

# create COVID2 var to indicate the time points starting from Mar 2020. 
datGynOUT1$COVID2 <- with(datGynOUT1, ifelse(time1 >= 2, 1, 0))

# create Post var; 1 should start at April 2020
datGynOUT1$post <- with(datGynOUT1, ifelse(time1 >= 3, (year-2020)*12+(month-3), 0))

```

```{r}
# aug 2021 number is particularly low, maybe due to data export error; use average of jul and sep 2021 to impute
datGynOUT1$count <- with(datGynOUT1, ifelse(year==2021 & month==8,
                                            round((datGynOUT1$count[datGynOUT1$year==2021 & datGynOUT1$month==7] +
                                         datGynOUT1$count[datGynOUT1$year==2021 & datGynOUT1$month==9]) / 2, 0), 
                                  count))

```


```{r}
# initial chart of # cases by month; just raw data (2015-2020)
ggplot(data = datGynOUT1, aes(x = time, y = count)) +
      geom_point() +
      geom_line() +
      ggtitle("Gynecology outpatient visits (Jan 2017 - May 2021)") +
      labs(x = "Time", y = "Number of gynecology outpatient visits") +
      scale_x_continuous(limits=c(24,76), breaks=c(24,36,48,60,72,76),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","May 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=68, y=20000, label="COVID-19",size=5) +
          expand_limits(y = 0) +
      goldentheme +
  theme(axis.text.x = element_text(size = 8.5)) -> pp.GynOUT
pp.GynOUT

```

### Analysis

$$ ln(E(y_t)) = \alpha_0 + \sum_{m=2}^{12} \beta_m Month + \beta_1 Time_t + \beta_2 COVID1 + \beta_3 COVID2_t + \beta_4 Post_t $$

```{r}
# adjust for monthly seasonality, January as the control，no offset for newborn population becuase 2021 data are not accurate
mod.GynOUT <- glm.nb(count ~ time1 + COVID1 + COVID2 + post 
                  #  + offset(log(newborn))
                  #  +feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec
                  + as.factor(month), 
                 data = subset(datGynOUT1, year>=2017))

#summary(mod.GynOUT)
lincom(mod.GynOUT)

# Extract the fitted values from this fitted model
datGynOUT1$fit <- NA
datGynOUT1$fit[datGynOUT1$year>=2017] <- fitted(mod.GynOUT)

# Use AR(1)
coef.GynOUT <- coeftest(mod.GynOUT, vcov=NeweyWest(mod.GynOUT,lag=1,prewhite = F))

# put coef, CI, and p-value for time1, COVID, post together
m.GynOUT <- matrix(, nrow = 4, ncol = 4)
m.GynOUT[,1]<-exp(coef.GynOUT[c(2:5),1])
m.GynOUT[,2]<-exp(coef.GynOUT[c(2:5),1]-1.96*coef.GynOUT[c(2:5),2])
m.GynOUT[,3]<-exp(coef.GynOUT[c(2:5),1]+1.96*coef.GynOUT[c(2:5),2])
m.GynOUT[,4]<-coef.GynOUT[c(2:5),4]
rownames(m.GynOUT) <- c("time1", "COVID1", "COVID2", "post")
colnames(m.GynOUT) <- c("exp(beta)", "95%low", "95%up", "p-value")
round(m.GynOUT, 5)

```


```{r}
# Counterfactual or expected Y values based on the original model
X0.GynOUT <- X.GynOUT <- model.matrix(mod.GynOUT)
X0.GynOUT[,3:5] <- 0
datGynOUT1$exp[datGynOUT1$year>=2017] <- as.vector(exp(X0.GynOUT%*%coef.GynOUT)) # save expected Y values to the monthly dataset

```

```{r}
####### 10,000 simulations
Sigma.GynOUT <- NeweyWest(mod.GynOUT,lag=1,prewhite=F)
set.seed(12)
beta.GynOUT <- mvrnorm(n = 10000, coef.GynOUT[,1], Sigma.GynOUT)

D.GynOUT <- Y_sum.GynOUT <- Y0_sum.GynOUT <- D_P.GynOUT <- RR.GynOUT <- NA

for (i in 1:10000){
  Y_sum.GynOUT[i] <- sum(as.vector(exp(X.GynOUT%*%beta.GynOUT[i,]))[38:53])
  Y0_sum.GynOUT[i] <- sum(as.vector(exp(X0.GynOUT%*%beta.GynOUT[i,]))[38:53])
  D.GynOUT[i] <- Y0_sum.GynOUT[i] - Y_sum.GynOUT[i]
  D_P.GynOUT[i] <- (Y0_sum.GynOUT[i] - Y_sum.GynOUT[i]) / Y0_sum.GynOUT[i]
  RR.GynOUT[i] <-  Y_sum.GynOUT[i] / Y0_sum.GynOUT[i]
}

# Fitted post
mean(Y_sum.GynOUT)
quantile(Y_sum.GynOUT, c(0.025,0.975)) #95% CI

# Counterfactual post
mean(Y0_sum.GynOUT)
quantile(Y0_sum.GynOUT, c(0.025,0.975)) 

# !!! Lost visits (observed - counterfactual/forecasted)
mean(D.GynOUT) 
quantile(D.GynOUT, c(0.025,0.975))
p.D.GynOUT <- length(D.GynOUT[D.GynOUT < 0]) / length(D.GynOUT) # p-value; if p=0, write it as p<0.0001
p.D.GynOUT

mean(D_P.GynOUT)  ## percentage of decrease
quantile(D_P.GynOUT, c(0.025,0.975))
p.D_P.GynOUT <- length(D_P.GynOUT[D_P.GynOUT < 0]) / length(D_P.GynOUT) # p-value
p.D_P.GynOUT

mean(RR.GynOUT)  ## RR
quantile(RR.GynOUT, c(0.025,0.975))

```

```{r}
### Chart for the model analysis
# CI around the line, just for the fitted line.
#agg_all$fit <- predict(m0,type="response")
datGynOUT1$fit_up[datGynOUT1$year>=2017] <- exp(predict(mod.GynOUT) + 1.96*predict(mod.GynOUT, se.fit=T)$se.fit)
datGynOUT1$fit_low[datGynOUT1$year>=2017] <- exp(predict(mod.GynOUT) - 1.96*predict(mod.GynOUT,se.fit=T)$se.fit)

# chart of observed versus expected
ggplot(data = datGynOUT1, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp, colour = "Expected")) +
        geom_line(aes(y = fit, colour = "Fitted")) + 
      geom_point(aes(y = count, colour = "Fitted")) +
      ggtitle("Gynecology outpatient visits: Observed versus expected (Jan 2017 - Nov 2021)") +
      labs(x = "Time", y = "Gynecology outpatient visits", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Dec 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=69, y=4000, label="COVID-19",size=5) +
#    scale_y_continuous(trans="log10") +
#          expand_limits(y = c(8000:26000)) +
      goldentheme +
  theme(axis.text.x = element_text(size = 9.5)) -> pp.GynOUT.ove
pp.GynOUT.ove

```


## Surgery

### Data & Descriptive

```{r}
# subset for surgery GYNO
datGynSUR <- subset(datOS, grepl("妇", datOS$keshi) & datOS$division=="surgery")

# combine counts if a monthly count was divided among more than 1 subdivisions of the obstetrics department
datGynSUR1 <- datGynSUR %>% group_by(year, month) %>% summarise_at(vars(count), sum)

# create Time var: first ever month in dataset = 0
datGynSUR1$time <- (datGynSUR1$year-2015)*12+(datGynSUR1$month-1)

# create Time1 var: 0 centered at Jan 2020 (1 time point before the pandemic actually happened in Feb 2020)
datGynSUR1$time1 <- datGynSUR1$time - 60

# create COVID1 var to indicate just Feb 2020 to make the drop in that month pop.
datGynSUR1$COVID1 <- with(datGynSUR1, ifelse(year == 2020 & month == 2, 1, 0))

# create COVID2 var to indicate the time points starting from Mar 2020. 
datGynSUR1$COVID2 <- with(datGynSUR1, ifelse(time1 >= 2, 1, 0))

# create Post var; 1 should start at April 2020
datGynSUR1$post <- with(datGynSUR1, ifelse(time1 >= 3, (year-2020)*12+(month-3), 0))

```

```{r}
# initial chart of # cases by month; just raw data (2015-2020)
ggplot(data = datGynSUR1, aes(x = time, y = count)) +
      geom_point() +
      geom_line() +
      ggtitle("Gynecology surgeries (Jan 2017 - May 2021)") +
      labs(x = "Time", y = "Number of gynecology surgeries") +
      scale_x_continuous(limits=c(24,76),breaks=c(24,36,48,60,72,76),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","May 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=68, y=100, label="COVID-19",size=5) +
          expand_limits(y = 0) +
      goldentheme +
  theme(axis.text.x = element_text(size = 8.5)) -> pp.GynSUR
pp.GynSUR

```

### Analysis

$$ ln(E(y_t)) = \alpha_0 + \sum_{m=2}^{12} \beta_m Month + \beta_1 Time_t + \beta_2 COVID1 + \beta_3 COVID2_t + \beta_4 Post_t $$

```{r}
# adjust for monthly seasonality, January as the control，no offset for newborn population becuase 2021 data are not accurate
mod.GynSUR <- glm.nb(count ~ time1 + COVID1 + COVID2 + post 
                  #  + offset(log(newborn))
                  #  +feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec
                  + as.factor(month), 
                 data = subset(datGynSUR1, year>=2017))

#summary(mod.GynSUR)
lincom(mod.GynSUR)

# Extract the fitted values from this fitted model
datGynSUR1$fit <- NA
datGynSUR1$fit[datGynSUR1$year>=2017] <- fitted(mod.GynSUR)

# Use AR(1)
coef.GynSUR <- coeftest(mod.GynSUR, vcov=NeweyWest(mod.GynSUR,lag=1,prewhite = F))

# put coef, CI, and p-value for time1, COVID, post together
m.GynSUR <- matrix(, nrow = 4, ncol = 4)
m.GynSUR[,1]<-exp(coef.GynSUR[c(2:5),1])
m.GynSUR[,2]<-exp(coef.GynSUR[c(2:5),1]-1.96*coef.GynSUR[c(2:5),2])
m.GynSUR[,3]<-exp(coef.GynSUR[c(2:5),1]+1.96*coef.GynSUR[c(2:5),2])
m.GynSUR[,4]<-coef.GynSUR[c(2:5),4]
rownames(m.GynSUR) <- c("time1", "COVID1", "COVID2", "post")
colnames(m.GynSUR) <- c("exp(beta)", "95%low", "95%up", "p-value")
round(m.GynSUR, 5)

```


```{r}
# Counterfactual or expected Y values based on the original model
X0.GynSUR <- X.GynSUR <- model.matrix(mod.GynSUR)
X0.GynSUR[,3:5] <- 0
datGynSUR1$exp[datGynSUR1$year>=2017] <- as.vector(exp(X0.GynSUR%*%coef.GynSUR)) # save expected Y values to the monthly dataset

```

```{r}
####### 10,000 simulations
Sigma.GynSUR <- NeweyWest(mod.GynSUR,lag=1,prewhite = F)
set.seed(12)
beta.GynSUR <- mvrnorm(n = 10000, coef.GynSUR[,1], Sigma.GynSUR)

D.GynSUR <- Y_sum.GynSUR <- Y0_sum.GynSUR <- D_P.GynSUR <- RR.GynSUR <- NA

for (i in 1:10000){
  Y_sum.GynSUR[i] <- sum(as.vector(exp(X.GynSUR%*%beta.GynSUR[i,]))[38:53])
  Y0_sum.GynSUR[i] <- sum(as.vector(exp(X0.GynSUR%*%beta.GynSUR[i,]))[38:53])
  D.GynSUR[i] <- Y0_sum.GynSUR[i] - Y_sum.GynSUR[i]
  D_P.GynSUR[i] <- (Y0_sum.GynSUR[i] - Y_sum.GynSUR[i]) / Y0_sum.GynSUR[i]
  RR.GynSUR[i] <-  Y_sum.GynSUR[i] / Y0_sum.GynSUR[i]
}

# Fitted post
mean(Y_sum.GynSUR)
quantile(Y_sum.GynSUR, c(0.025,0.975)) #95% CI

# Counterfactual post
mean(Y0_sum.GynSUR)
quantile(Y0_sum.GynSUR, c(0.025,0.975)) 

# !!! Lost visits (observed - counterfactual/forecasted)
mean(D.GynSUR) 
quantile(D.GynSUR, c(0.025,0.975))
p.D.GynSUR <- length(D.GynSUR[D.GynSUR < 0]) / length(D.GynSUR) # p-value; if p=0, write it as p<0.0001
p.D.GynSUR

mean(D_P.GynSUR)  ## percentage of decrease
quantile(D_P.GynSUR, c(0.025,0.975))
p.D_P.GynSUR <- length(D_P.GynSUR[D_P.GynSUR < 0]) / length(D_P.GynSUR) # p-value
p.D_P.GynSUR

mean(RR.GynSUR)  ## RR
quantile(RR.GynSUR, c(0.025,0.975))

```

```{r}
### Chart for the model analysis
# CI around the line, just for the fitted line.
#agg_all$fit <- predict(m0,type="response")
datGynSUR1$fit_up[datGynSUR1$year>=2017] <- exp(predict(mod.GynSUR) + 1.96*predict(mod.GynSUR, se.fit=T)$se.fit)
datGynSUR1$fit_low[datGynSUR1$year>=2017] <- exp(predict(mod.GynSUR) - 1.96*predict(mod.GynSUR,se.fit=T)$se.fit)

# chart of observed versus expected
ggplot(data = datGynSUR1, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp, colour = "Expected")) +
        geom_line(aes(y = fit, colour = "Fitted")) + 
      geom_point(aes(y = count, colour = "Fitted")) +
      ggtitle("Gynecology surgeries: Observed versus expected (Jan 2017 - Nov 2021)") +
      labs(x = "Time", y = "Gynecology surgeries", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Dec 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=68, y=150, label="COVID-19",size=5) +
#          expand_limits(y = c(8000:26000)) +
      goldentheme +
  theme(axis.text.x = element_text(size = 9.5)) -> pp.GynSUR.ove
pp.GynSUR.ove

```


## Inpatient

### Date & Descriptive

```{r, echo=T}
# read in data
datGYN <- read_csv("Data/allgyn.csv")

# create Time var: first ever month in dataset = 0
datGYN$time <- (datGYN$year-2015)*12+(datGYN$month-1)

# create Time1 var: 0 centered at Jan 2020 (1 time point before the pandemic actually happened in Feb 2020)
datGYN$time1 <- datGYN$time - 60

# create COVID1 var to indicate just Feb 2020 to make the drop in that month pop.
datGYN$COVID1 <- with(datGYN, ifelse(year == 2020 & month == 2, 1, 0))

# create COVID2 var to indicate the time points starting from Mar 2020. 
datGYN$COVID2 <- with(datGYN, ifelse(time1 >= 2, 1, 0))

# create Post var; 1 should start at April 2020
datGYN$post <- with(datGYN, ifelse(time1 >= 3, (year-2020)*12+(month-3), 0))

```


```{r}
# Summarize monthly counts
datGYN %>% 
  group_by(time, year, month, time1, COVID1, COVID2, post) %>% 
  count(time) -> datGynIN

# june 2021 is missing. need to add a line and use average of may and aug 2021 to impute
datGynIN$year[84] <- 2021
datGynIN$month[84] <- 6
datGynIN$time[84] <- (datGynIN$year[84]-2015)*12+(datGynIN$month[84]-1)
datGynIN$time1[84] <- datGynIN$time[84] - 60
datGynIN$COVID1[84] <- with(datGynIN, ifelse(year[84] == 2020 & month[84] == 2, 1, 0))
datGynIN$COVID2[84] <- with(datGynIN, ifelse(time1[84] >= 2, 1, 0))
datGynIN$post[84] <- with(datGynIN, ifelse(time1[84] >= 3, (year[84]-2020)*12+(month[84]-3), 0))

# remove the last row for summarizing NA date/time info
datGynIN <- datGynIN[-c(85),]

# july 2021 is exceptionally low because of move into new hospital
## use average of may and aug 2021 to impute for june and july 2021
datGynIN$n <- with(datGynIN, ifelse(year==2021 & month==6 | year==2021 & month==7,
                                    round((datGynIN$n[datGynIN$year==2021 & datGynIN$month==5] +
                                         datGynIN$n[datGynIN$year==2021 & datGynIN$month==8]) / 2, 0), 
                                  n))
```

```{r}
### GYNECOLOGY
# initial chart of # cases by month; just raw data (2015-2020)
ggplot(data = datGynIN, aes(x = time, y = n)) +
      geom_point() +
      geom_line() +
      ggtitle("Gynecology inpatient admissions (Jan 2017 - May 2021)") +
      labs(x = "Time", y = "Number of gynecology admissions") +
      scale_x_continuous(limits=c(24,76),breaks=c(24,36,48,60,72,76),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","May 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=68, y=1200, label="COVID-19",size=5) +
          expand_limits(y = 0) +
      goldentheme +
  theme(axis.text.x = element_text(size = 8.5)) -> pp.GynIN
pp.GynIN

```

### Analysis

$$ ln(E(y_t)) = \alpha_0 + \sum_{m=2}^{12} \beta_m Month + \beta_1 Time_t + \beta_2 COVID1 + \beta_3 COVID2_t + \beta_4 Post_t $$

```{r}
# adjust for monthly seasonality, January as the control，no offset for newborn population becuase 2021 data are not accurate
mod.GynIN <- glm.nb(n ~ time1 + COVID1 + COVID2 + post 
                  #  + offset(log(newborn))
                  #  +feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec
                  + as.factor(month), 
                 data = subset(datGynIN, year>=2017))

#summary(mod.GynIN)
lincom(mod.GynIN)

# Extract the fitted values from this fitted model
datGynIN$fit <- NA
datGynIN$fit[datGynIN$year>=2017] <- fitted(mod.GynIN)

# Use AR(1)
coef.GynIN <- coeftest(mod.GynIN, vcov=NeweyWest(mod.GynIN,lag=1,prewhite = F))

# put coef, CI, and p-value for time1, COVID, post together
m.GynIN <- matrix(, nrow = 4, ncol = 4)
m.GynIN[,1]<-exp(coef.GynIN[c(2:5),1])
m.GynIN[,2]<-exp(coef.GynIN[c(2:5),1]-1.96*coef.GynIN[c(2:5),2])
m.GynIN[,3]<-exp(coef.GynIN[c(2:5),1]+1.96*coef.GynIN[c(2:5),2])
m.GynIN[,4]<-coef.GynIN[c(2:5),4]
rownames(m.GynIN) <- c("time1", "COVID1", "COVID2", "post")
colnames(m.GynIN) <- c("exp(beta)", "95%low", "95%up", "p-value")
round(m.GynIN, 5)

```


```{r}
# Counterfactual or expected Y values based on the original model
X0.GynIN <- X.GynIN <- model.matrix(mod.GynIN)
X0.GynIN[,3:5] <- 0
datGynIN$exp[datGynIN$year>=2017] <- as.vector(exp(X0.GynIN%*%coef.GynIN)) # save expected Y values to the monthly dataset

```

```{r}
####### 10,000 simulations
Sigma.GynIN <- NeweyWest(mod.GynIN,lag=1,prewhite = F)
set.seed(12)
beta.GynIN <- mvrnorm(n = 10000, coef.GynIN[,1], Sigma.GynIN)

D.GynIN <- Y_sum.GynIN <- Y0_sum.GynIN <- D_P.GynIN <- RR.GynIN <- NA

for (i in 1:10000){
  Y_sum.GynIN[i] <- sum(as.vector(exp(X.GynIN%*%beta.GynIN[i,]))[38:53])
  Y0_sum.GynIN[i] <- sum(as.vector(exp(X0.GynIN%*%beta.GynIN[i,]))[38:53])
  D.GynIN[i] <- Y0_sum.GynIN[i] - Y_sum.GynIN[i]
  D_P.GynIN[i] <- (Y0_sum.GynIN[i] - Y_sum.GynIN[i]) / Y0_sum.GynIN[i]
  RR.GynIN[i] <-  Y_sum.GynIN[i] / Y0_sum.GynIN[i]
}

# Fitted post
mean(Y_sum.GynIN)
quantile(Y_sum.GynIN, c(0.025,0.975)) #95% CI

# Counterfactual post
mean(Y0_sum.GynIN)
quantile(Y0_sum.GynIN, c(0.025,0.975)) 

# !!! Lost visits (observed - counterfactual/forecasted)
mean(D.GynIN) 
quantile(D.GynIN, c(0.025,0.975))
p.D.GynIN <- length(D.GynIN[D.GynIN < 0]) / length(D.GynIN) # p-value; if p=0, write it as p<0.0001
p.D.GynIN

mean(D_P.GynIN)  ## percentage of decrease
quantile(D_P.GynIN, c(0.025,0.975))
p.D_P.GynIN <- length(D_P.GynIN[D_P.GynIN < 0]) / length(D_P.GynIN) # p-value
p.D_P.GynIN

mean(RR.GynIN)  ## RR
quantile(RR.GynIN, c(0.025,0.975))

```

```{r}
### Chart for the model analysis
# CI around the line, just for the fitted line.
#agg_all$fit <- predict(m0,type="response")
datGynIN$fit_up[datGynIN$year>=2017] <- exp(predict(mod.GynIN) + 1.96*predict(mod.GynIN, se.fit=T)$se.fit)
datGynIN$fit_low[datGynIN$year>=2017] <- exp(predict(mod.GynIN) - 1.96*predict(mod.GynIN,se.fit=T)$se.fit)

# chart of observed versus expected
ggplot(data = datGynIN, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp, colour = "Expected")) +
        geom_line(aes(y = fit, colour = "Fitted")) + 
      geom_point(aes(y = n, colour = "Fitted")) +
      ggtitle("Gynecology inpatient admissions: \nObserved versus expected (Jan 2017 - May 2021)") +
      labs(x = "Time", y = "Gynecology admissions", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Dec 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=68.5, y=200, label="COVID-19",size=5) +
#          expand_limits(y = c(8000:26000)) +
      goldentheme +
  theme(axis.text.x = element_text(size = 9.5)) -> pp.GynIN.ove
pp.GynIN.ove

```

### Plot all 3 charts together

```{r}
# outpatient
ggplot(data = datGynOUT1, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp), color = "#0072B2", linetype = "dashed") +
        geom_line(aes(y = fit), color = "#0072B2",) + 
      geom_point(aes(y = count), color = "#0072B2", size = 1) +
      ggtitle("A. Outpatient visits") +
      labs(x = "", y = "", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                         labels=addline_format(c("","","","","",""))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
       scale_y_continuous(trans="log10") +
      goldentheme_facet +
  theme(axis.title.y=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none",
        plot.title = element_text(face = "bold")) -> pp.GynOUT.ove1
pp.GynOUT.ove1

# surgeries
ggplot(data = datGynSUR1, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp), color = "#000000", linetype = "dashed") +
        geom_line(aes(y = fit), color = "#000000") + 
      geom_point(aes(y = count), color = "#000000", size = 1) +
      ggtitle("B. Surgeries") +
      labs(x = "", y = "", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                         labels=addline_format(c("","","","","",""))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
          expand_limits(y = c(500:1200)) +
         scale_y_continuous(trans="log10") +
      goldentheme_facet +
  theme(axis.title.y=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none",
        plot.title = element_text(face = "bold")) -> pp.GynSUR.ove1
pp.GynSUR.ove1

#inpatient
ggplot(data = datGynIN, aes(x = time)) +
    geom_ribbon(aes(ymin = fit_low, ymax = fit_up), alpha = 0.15) +
  geom_line(aes(y = exp), color = "#D55E00", linetype = "dashed") +
        geom_line(aes(y = fit), color = "#D55E00") + 
      geom_point(aes(y = n), color = "#D55E00", size = 1) +
      ggtitle("C. Inpatient admissions") +
      labs(x = "Time", y = "", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Dec 2021"))) +
      geom_vline(xintercept = 61, linetype = "dotted") +
    annotate(geom="text", x=67.5, y=1300, label="COVID-19",size=4) +
#          expand_limits(y = c(1000:2000)) +
         scale_y_continuous(trans="log10") +
      goldentheme_facet +
  theme(axis.text.x = element_text(size = 10),
        axis.title.y=element_blank(),
        legend.position = "none",
        plot.title = element_text(face = "bold")) -> pp.GynIN.ove1
pp.GynIN.ove1

```

```{r}
threegyn <- plot_grid(pp.GynOUT.ove1, pp.GynSUR.ove1, pp.GynIN.ove1, nrow = 3,
          rel_heights = c(1, 1, 1.15))

ggsave("~/Documents/RESEARCH/Shandong/Manuscripts/Figures/Gyn3in1.pdf", width = 5, height = 8)

threegyn

```

```{r}
# all 3 lines in one chart
ggplot() +
        geom_vline(xintercept = 61, linetype = "dotted") +
  geom_ribbon(data = datGynOUT1, aes(x = time, ymin = fit_low, ymax = fit_up), fill = "#0072B2", alpha = 0.15) +
  geom_ribbon(data = datGynSUR1, aes(x = time, ymin = fit_low, ymax = fit_up), fill = "#000000", alpha = 0.15) +
  geom_ribbon(data = datGynIN, aes(x = time, ymin = fit_low, ymax = fit_up), fill = "#D55E00", alpha = 0.15) +
      geom_line(data = datGynOUT1, aes(x = time, y = exp), color = "#0072B2", linetype = "dashed") +
      geom_line(data = datGynSUR1, aes(x = time, y = exp), color = "#000000", linetype = "dashed") +
      geom_line(data = datGynIN, aes(x = time, y = exp), color = "#D55E00", linetype = "dashed") +
        geom_line(data = datGynOUT1, aes(x = time, y = fit), color = "#0072B2",) + 
        geom_line(data = datGynSUR1, aes(x = time, y = fit), color = "#000000") + 
        geom_line(data = datGynIN, aes(x = time, y = fit), color = "#D55E00") + 
      geom_point(data = datGynOUT1, aes(x = time, y = count), color = "#0072B2", size = 1) +
      geom_point(data = datGynSUR1, aes(x = time, y = count), color = "#000000", size = 1) +
      geom_point(data = datGynIN, aes(x = time, y = n), color = "#D55E00", size = 1) +
      ggtitle("") +
      labs(x = "Time", y = "", color = "") +
      scale_x_continuous(limits=c(24,83), breaks=c(24,36,48,60,72,83),
                       labels=addline_format(c("Jan 2017","Jan 2018","Jan 2019","Jan 2020","Jan 2021","Dec 2021"))) +
      annotate(geom="text", x=67, y=3500, label="COVID-19",size=5) +
          annotate(geom="text", x=29, y=8200, label="Outpatient visits",size=4, color = "#0072B2") +
        annotate(geom="text", x=30.5, y=940, label="Inpatient admissions",size=4, color = "#D55E00") +
          annotate(geom="text", x=27, y=330, label="Surgeries",size=4, color = "#000000") +
            expand_limits(y = c(100:30000)) +
       scale_y_continuous(trans="log10") +
      goldentheme +
  theme(axis.title.y=element_blank(),
        legend.position = "none",
        plot.title = element_blank()) -> pp.Gyn.all.ove

ggsave("~/Documents/RESEARCH/Shandong/Manuscripts/Figures/Gyn3in1_v2.pdf", width = 8, height = 5)

pp.Gyn.all.ove

```


***

# Appendix

```{r,ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
```